### 2023-03-27 paper/admin/cover_letter.tex
lots of bug fixes (no results are affected; all bugs were related to running the code in a fresh environment)

\documentclass[10pt,stdletter,orderfromtodate,sigleft]{newlfm}
\usepackage{blindtext, xfrac, animate, hyperref, pxfonts}

  \setlength{\voffset}{0in}

\newlfmP{dateskipbefore=0pt}
\newlfmP{sigsize=20pt}
\newlfmP{sigskipbefore=10pt}

\newlfmP{Headlinewd=0pt,Footlinewd=0pt}

\namefrom{Jeremy R. Manning, Ph.D.}
\addrfrom{
	Dartmouth College\\
  Department of Psychological \& Brain Sciences\\
  HB 6207 Moore Hall\\
	Hanover, NH  03755}

\addrto{}
\dateset{\today}

\greetto{To the editors of \textit{Scientific Reports}:}

\closeline{Sincerely,}

\begin{document}
\begin{newlfm}

We have enclosed our manuscript entitled \textit{Category-based and
location-based volitional covert attention affect memory at different
timescales}, which we wish to submit for publication as an Article in
\textit{Scientific Reports}.

Our manuscript reports our findings from an experiment designed to study how
different aspects of covert attention affect memory. Our experiment comprised
two conditions that asked participants to \textit{sustain} or \textit{vary} the
focus of their covert attention, respectively, with respect to each of a series
of stimuli. Each stimulus comprised a pair of images (one on the left of the
screen and the other on the right), and each image comprised an equal blend of
a photograph of a face and a photograph of a scene. During each stimulus
presentation, we instructed participants to focus in on either the face or
scene component of either the left or right image \textit{without moving their
eyes}. (This is quite difficult to do naively, so we trained participants to do
this quickly and at will using a practice session prior to the main
experiment.) We used eye-tracking data to focus in on trials where participants
were specifically varying their focus of \textit{covert} attention (i.e., with
no change in where they were looking), as opposed to simply moving their eyes
to look at the to-be-attended images. We then administered recognition memory
tests that asked participants to rate the ``familiarity'' of attended,
unattended, and novel images.

As we anticipated, participants in both conditions remembered stimuli they
attended to better than unattended stimuli, and they rated both attended and
unattended stimuli they had seen as more ``familiar'' than novel images, during
a recognition memory test. We also found some interesting differences across
the two experimental conditions pertaining to how participants remembered
``partially attended'' stimuli. When participants varied the focus of their
attention more often, their memory was more influenced by location-based
attention (regardless of which image category they were attending). In
contrast, when participants sustained the focus of their attention over longer
intervals, their memory was more influenced by category-based attention
(regardless of which location they were attending). Our findings highlight a
dissociation between how location-based and category-based attention affect
memory encoding (and subsequent recognition). Location-based attention appears
to affect memory encoding on relatively short timescales by enhancing encoding
for stimuli at the attended location. On the other hand, category-based
attention appears to affect memory encoding on longer timescales by
\textit{suppressing} stimuli from the unattended category.

The impact of attention on memory is of central importance to a wide
variety of scientifically and socially current issues.  For example,
the interplay between attention and memory affects how we remember our
ongoing experiences in everyday life, how students learn in the
classroom (and which teaching strategies might be most effective), how
(or whether) we recognize a face we notice in a crowd and
happen to encounter later, and so on.  These issues also relate to
attention disorders (e.g., ADD, ADHD) and memory disorders
(e.g., age-related memory impairment, Alzheimer's, etc.).  Discovering
the mechanisms that underlie how our experiences are encoded into
memories can lead to new tools for identifying when those mechanisms
are not operating within their expected parameters.

We have suggested the following scientists as expert reviewers: Marvin Chun
(Marvin.Chun@Yale.edu), Brice Kuhl (bkuhl@uoregon.edu), Joel Voss
(Joel-Voss@Northwestern.edu), Edward Vogel (Ed\-Vogel@UChi\-ca\-go.edu), and
Judith Fan (jefan@ucsd.edu). We have also published and documented all of our
experimental stimuli, code, and data, along with the code used to generate the
figures and analyses we report in our manuscript. These may be found at
\href{https://github.com/ContextLab/attention-memory-task}{\texttt{https://github.com/ContextLab/attention-memory-task}}.

Thank you for considering our manuscript for publication in
Psychonomic Bulletin \& Review.

\end{newlfm}
\end{document}